Constraints on quasi--normal-mode frequencies with LIGO-Virgo binary--black-hole observations

### 2021-03-26 papers/intro_paper/results_synthetic.tex
updated draft to address Richard's comments

To avoid possible systematic biases in our parameter-estimation analysis
due to error in waveform modeling, we use the GR version of the same waveform,
$\SEOB$ (i.e., without allowing for deviations in the QNM parameters) to
simulate our GW signal. And to avoid systematic biases due to noise,
we use an averaged (zero-noise) realization of the noise. \footnote{A detailed
study on noise systematics for one of the GW events is presented in
Appendix~\ref{sec:noise_systematics}.}  As in the case of the actual
detections of GW150914 and GW190521, we consider a two-detector LIGO network at
Hanford and Livingston, having identical PSDs \textcolor{red}{RB:did we use Virgo? here it seems not, but above we say we used Virgo}. The distance to the two 
synthetic events is rescaled such that the SNR in the detector network
is the same as the actual events (i.e., 24 for GW150914 and 14 \textcolor{red}{RB: $\rho_{\rm IMR}$ for the GW190521-like injection shown in the table is larger than this...}
for GW190521). Since mearly--equal-mass binaries like GW150914 and
GW190521 observed at moderately high SNRs are not expected to have a
loud ringdown signal, we restrict ourselves to estimating the
frequency and damping time of only one QNM $(\ell m) = (2,2)$ (i.e.,
$\{\df{220},\dtau{220}\}$), while fixing the other QNM frequencies to
their GR values.

@@

We additionally investigate the effects of erroneously assuming that
an underlying non-GR signal can be well-described by a GR one. We do
this by estimating the parameters of our non-GR signals using the GR
waveform model $\SEOB$ instead of the parameterized $\pSEOB$.  The
resulting one- and two-dimensional posteriors are shown in the right
panels of Fig.~\ref{fig:simulated_signal_nonGR} by red curves for the
GW150914-like (top) and GW190521-like (bottom) signals
respectively. For both signals, we find the $\SEOB$ estimates are
markedly biased with respect to the $\pSEOB$ estimates. We also notice that the
results are distinctly different for the two events. For the
GW150914-like non-GR signal, the measurements of $(\fgr{220},
\taugr{220})$ (top right panel in
Fig.~\ref{fig:simulated_signal_nonGR}) are consistent with the
$(\fngr{220}, \taungr{220})$ measurements for a signal with \emph{no
  deviations} from GR (top right panel in
Fig.~\ref{fig:simulated_signal_GR}). In other words, if the actual
signal had deviations from GR as large as the $10\%$, the analysis
with the GR signal $\SEOB$ would likely have reported \emph{no}
deviation from the GR prediction \textcolor{red}{RB: not sure what this means, since a GR waveform by itself cannot detect a non-GR deviation}. However in the case of the
GW190521-like non-GR signal, a simple GR analysis of the non-GR signal
would have yielded measurements distinctly different from either of
the two parameterized estimates: with and without deviations\textcolor{red}{RB: I don't understand this previous sentence. Discuss with Abhirup}.

Finally, we investigate the impact on the estimation of the GR parameters $\bxigr$ as we change the
magnitude of deviation in the non-GR parameters. We restrict ourselves 
to a GW150914-like event, and for comparison, add a synthetic signal
with $\df{220} = \dtau{220} = 0.5$, or 50 \% deviation, alongside the
10\% non-GR signal and a signal with no deviation (essentially GR)
mentioned above. We summarise the results in
Fig.~\ref{fig:gr_ngr_comparison}. The left (right) panels show the
posterior probability distributions of the three signals using the GR,
$\SEOB$ (non-GR, $\pSEOB$) waveform model. The first difference we
note is the $\SEOB$ recoveries yield biased estimates when the
underlying signal is non-GR, while the $\pSEOB$ recoveries do
not. Furthermore, as we increase the deviations, while the $\SEOB$
recoveries expectedly get more biased, the $\pSEOB$ measurements are
robust in consistently recovering the injected value. Since the
measurement of the frequency and damping time in the $\pSEOB$ model
incorporate information from the masses and spins through the NR
fitting formulae (as shown in
Eqs.~(\ref{eq:nongr_freqs_a})---(\ref{eq:nongr_freqs_b})), one might have
expected correlations between the mass, spins and the QNM frequencies,
leading to biased estimates in the masses and spins. We confirm here
that is not the case. \comment{AB: I don't see how from the figure I can  
conclude that there are no correlations between masses/spins and QNMs.}

@@

We summarize our results in Fig.~\ref{fig:nohair_sxs}.  Given the
injection parameters, the predicted values of the $(2, 2)$ and
$(3, 3)$ frequency and damping time are \macro{(169.45 Hz, 4.68
  ms)} and \macro{(271.21 Hz, 4.50 ms)} respectively. The left panel
of Fig.~\ref{fig:nohair_sxs} shows that the 2D posteriors
on the $(2, 2)$ and $(3, 3)$ QNMs are consistent with the
predictions for a BBH merger in GR, indicated by the black plus sign.
Using fitting formulae provided in Ref.~\cite{Berti:2005ys}, specifically,
Eq.~(2.1), Eq.~(E1) and Eq.~(E3) and Tables VIII and IX for the fitting coefficients,
we infer the 2D posterior probability distribution on the
final mass and spin \rb{as predicted independently by} the $(2, 2)$ (blue) and $(3, 3)$ (red)
QNMs in the right panel of Fig.~\ref{fig:nohair_sxs}. The two
independent estimates are consistent with each other and correspond to
a unique mass and spin for the remnant BH \macro{(83.08 $\Mo$, 0.37)}
indicated by the plus sign. As a consequence, this may be considered
as a test of the no-hair conjecture. For most of the events observed
so far, the power in the $(3, 3)$ has not been sufficient to
measure it along with the $(2, 2)$, or in fact, in its
place \textcolor{red}{RB:"in its place" is weird, since the (2,2) is always louder than the (3,3) mode (unless you're thinking about very high mass BHs for which the (2,2) mode would be out-of-band, but the (3,3) still in band. But in any case, I would finish the sentence at "has not been sufficient to measure it along with the $(2, 2)$"}. However, it might also be possible to combine information from
multiple observation over the coming few years to obtain meaningful
constraints on the $(3, 3)$ and other sub-dominant QNMs. \textcolor{red}{reminder: citations needed here}

>>>>

To avoid possible systematic biases in our parameter-estimation analysis
due to error in waveform modeling, we use the GR version of the same waveform,
$\SEOB$ (i.e., without allowing for deviations in the QNM parameters) to
simulate our GW signal. And to avoid systematic biases due to noise,
we use an averaged (zero-noise) realization of the noise. \footnote{A detailed
study on noise systematics for one of the GW events is presented in
Appendix~\ref{sec:noise_systematics}.}  Since mearly--equal-mass binaries like GW150914 and
GW190521 observed at moderately high SNRs are not expected to have a
loud ringdown signal, we restrict ourselves to estimating the
frequency and damping time of only one QNM $(\ell m) = (2,2)$ (i.e.,
$\{\df{220},\dtau{220}\}$), while fixing the other QNM frequencies to
their GR values.

@@

We additionally investigate the effects of erroneously assuming that
an underlying non-GR signal can be well-described by a GR one. We do
this by estimating the parameters of our non-GR signals using the GR
waveform model $\SEOB$ instead of the parameterized $\pSEOB$.  The
resulting one- and two-dimensional posteriors are shown in the right
panels of Fig.~\ref{fig:simulated_signal_nonGR} by red curves for the
GW150914-like (top) and GW190521-like (bottom) signals
respectively. For both signals, we find the $\SEOB$ estimates are
markedly biased with respect to the $\pSEOB$ estimates.  We investigate 
this impact on the estimation of the GR parameters $\bxigr$ by changing the
magnitude of deviation in the non-GR parameters more in Fig.~\ref{fig:gr_ngr_comparison}. We restrict ourselves 
to a GW150914-like event, and for comparison, add a synthetic signal
with $\df{220} = \dtau{220} = 0.5$, or 50 \% deviation, alongside the
10\% non-GR signal and a signal with no deviation (essentially GR)
mentioned above. The left (right) panels show the
posterior probability distributions of the three signals using the GR,
$\SEOB$ (non-GR, $\pSEOB$) waveform model. The first difference we
note is the $\SEOB$ recoveries yield biased estimates when the
underlying signal is non-GR, while the $\pSEOB$ recoveries do
not. Furthermore, as we increase the deviations, while the $\SEOB$
recoveries expectedly get more biased, the $\pSEOB$ measurements are
robust in consistently recovering the injected value. \abhi{This gives us 
confidence that the $\pSEOB$ model can accurately measure QNM 
properties without biasing measurements of inspiral quantities, like masses and spins.}

@@

We summarize our results in Fig.~\ref{fig:nohair_sxs}.  Given the
injection parameters, the predicted values of the $(2, 2)$ and
$(3, 3)$ frequency and damping time are \macro{(169.45 Hz, 4.68
  ms)} and \macro{(271.21 Hz, 4.50 ms)} respectively. The left panel
of Fig.~\ref{fig:nohair_sxs} shows that the 2D posteriors
on the $(2, 2)$ and $(3, 3)$ QNMs are consistent with the
predictions for a BBH merger in GR, indicated by the black plus sign.
Using fitting formulae provided in Ref.~\cite{Berti:2005ys}, specifically,
Eq.~(2.1), Eq.~(E1) and Eq.~(E3) and Tables VIII and IX for the fitting coefficients,
we infer the 2D posterior probability distribution on the
final mass and spin \rb{as predicted independently by} the $(2, 2)$ (blue) and $(3, 3)$ (red)
QNMs in the right panel of Fig.~\ref{fig:nohair_sxs}. The two
independent estimates are consistent with each other and correspond to
a unique mass and spin for the remnant BH \macro{(83.08 $\Mo$, 0.37)}
indicated by the plus sign. As a consequence, this may be considered
as a test of the no-hair conjecture. For most of the events observed
so far, the power in the $(3, 3)$ has not been sufficient to
measure it along with the $(2, 2)$. However, it might also be possible to combine information from
multiple observation over the coming few years to obtain meaningful
constraints on the $(3, 3)$ and other sub-dominant QNMs~\cite{Gossan:2011ha,Meidam:2014jpa,Yang:2017zxs}.

### 2021-03-26 papers/intro_paper/table_injection_values.tex
changes in Table II, Sec. IV and V
minor changes
minor updates
updated draft to address Richard's comments

\begin{table}[h!]
\begin{center}
\begin{tabular}{ c|c|c|c|c|c|c|c }

 \ab{synthetic signal} & \makecell{$m_{\rm 1\,det}$ \\$(\Mo)$} &  \makecell{$m_{\rm 2\,det}$ \\ $(\Mo)$} & $\chi_{1}$ & $\chi_{2}$ & $\rho_\text{IMR}$ & $\rho_\text{insp}$ & $\rho_\text{postinsp}$ \\
 \hline
 GW150914-like & 39 & 31 & 0.0 & 0.0 & 25 & 22 & 12 \\
 GW190521-like & 150 & 120 & 0.02 & -0.39 & 20 & 8 & 18 \\
 SXS:BBH:0166 & 72 & 12  & 0.0 & 0.0 & 71 & 58 & 41 \\

\end{tabular}
\caption{Parameters of the synthetic-signal injections, chosen to be similar to the actual GW events indicated in the first column (first two rows). The parameters $(m_{\rm 1\,det},m_{\rm 2 \,det})$ are the detector-frame masses of the primary and secondary BHs, respectively. The third row indicates the parameters of the SXS BBH waveform used in Sec.~\ref{ssec:nohairtheorem}. The quantities $\rho_\text{IMR}$, $\rho_\text{insp}$ and $\rho_\text{postinsp}$ are the SNR of the full IMR signal, SNR upto a certain cutoff frequency, and SNR after the cutoff frequency respectively. In each case, the cutoff frequency is assumed to be the frequency at the innermost circular stable orbit (ISCO) corresponding to the remnant Kerr BH.}
\label{tab:injection_values}
\end{center}
\end{table}

>>>>

\begin{table}[h!]
\begin{center}
\begin{tabular}{ c|c|c|c|c|c|c|c|c }

Injection &  Network & \makecell{$m_{\rm 1\,det}$ \\$(\Mo)$} &  \makecell{$m_{\rm 2\,det}$ \\ $(\Mo)$} & $\chi_{1}$ & $\chi_{2}$ & $\rho_\text{IMR}$ & $\rho_\text{insp}$ & $\rho_\text{postinsp}$ \\
 \hline
 GW150914-like & HL &39 & 31 & 0.0 & 0.0 & 25 & 22 & 12 \\
 GW190521-like & HL & 150 & 120 & 0.02 & -0.39 & 20 & 8 & 18 \\
 SXS:BBH:0166 & HLV &72 & 12  & 0.0 & 0.0 & 71 & 58 & 41 \\

\end{tabular}
\caption{Parameters of the synthetic-signal injections, chosen to be similar to the actual GW events indicated in the first column (first two rows). The parameters $(m_{\rm 1 \,det}, m_{\rm 2 \,det})$ are the detector-frame masses of the primary and secondary BHs, respectively. The third row indicates the parameters of the SXS BBH waveform used in Sec.~\ref{ssec:nohairtheorem}. The second column refers to the detector-network used, with H,L,V, referring to LIGO-Hanford, LIGO-Livingston and Virgo, respectively. The quantities $\rho_\text{IMR}$, $\rho_\text{insp}$ and $\rho_\text{postinsp}$ are the SNR of the full IMR signal, SNR upto a certain cutoff frequency, and SNR after the cutoff frequency respectively. In each case, the cutoff frequency is assumed to be the frequency at the innermost circular stable orbit (ISCO) corresponding to the remnant Kerr BH.}
\label{tab:injection_values}
\end{center}
\end{table}
